### 2023-04-04 paper-reviews/papers-2023.tex
Add dynamic warp formation paper

    \cite{Burns:IntelHTTValidation:2002}:

@@

    The authors discovered that GPU programs contain significant portion of scalar instructions. 30\% for the GPU selected benchmarks in the paper. The papers states that scalar instructions are typically executed as SIMD with all vector elements having the same values. Thus, the same computation is performed redundantly. One of the solution for the problem is to run the calculations once and than broadcast the results in order to reduce power consumption on unnecessary operations. The authors suggested to statically detect scalar operations during compilations. One negative effect of this is reduced utilization of SIMD lanes. To improve utilization the  paper proposed scalar waving --- grouping of same scalar instructions processing different scalars into a single SIMD instruction. This grouping is suggest to be performed during runtime through by GPU itself. Further, the authors developed simultaneous scalar and SIMD group waving --- grouping and a SIMD instructions and scalar ware. GPU needs to be able to execute two different operations as a part of SIMD instruction in this case. The authors claim to achieve up to 25\% percent performance improvement for the selected benchmarks. The performance studies are performed on \href{https://github.com/gpgpu-sim}{GPGPU-Sim} simulator.
\end{itemize}

>>>>

    \item \cite{Burns:IntelHTTValidation:2002}:

@@

    The authors discovered that GPU programs contain significant portion of scalar instructions. 30\% for the GPU selected benchmarks in the paper. The papers states that scalar instructions are typically executed as SIMD with all vector elements having the same values. Thus, the same computation is performed redundantly. One of the solution for the problem is to run the calculations once and than broadcast the results in order to reduce power consumption on unnecessary operations. The authors suggested to statically detect scalar operations during compilations. One negative effect of this is reduced utilization of SIMD lanes. To improve utilization the  paper proposed scalar waving --- grouping of same scalar instructions processing different scalars into a single SIMD instruction. This grouping is suggest to be performed during runtime through by GPU itself. Further, the authors developed simultaneous scalar and SIMD group waving --- grouping and a SIMD instructions and scalar ware. GPU needs to be able to execute two different operations as a part of SIMD instruction in this case. The authors claim to achieve up to 25\% percent performance improvement for the selected benchmarks. The performance studies are performed on \href{https://github.com/gpgpu-sim}{GPGPU-Sim} simulator.

    \item \cite{Fung:DynamicWrapGPUControlFlow:2007}:

    The paper is focused on GPGPU control flow problems that represent general control flow issues with SIMD --- what should be done if control flow diverges for SIMD lines. The paper describes several possible solutions for the problem: scalar execution (more typical for CPUs), reconvergence at the immediate post-dominator, dynamic warp formation that is supposed to improve SIMD pipeline utilization. The authors show that reconvergence at the immediate post-dominator is not always the best choice. They propose an implementation of dynamic warp formation together with several heuristics that are supposed to define when the formed warps should be pushed for execution. The authors claim that dynamic warp formation improves performance by 20.7\% on average compared to reconvergence at the immediate post-dominator. They estimated that area increase because of the proposed hardware changes to be around 4.7\%.
\end{itemize}
